---
title: "Visualizing Impact of Great Recession"
author: "Deeksha Patil"
date: "4/10/2022"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
      vertical_layout: fill
---

```{r Libraries, include=FALSE, message=FALSE}

library(tidyverse)
library(DT)
library(Hmisc)
library(lubridate)
library(plotly)
library(gridExtra)
library(maps)
library(shiny)
library(shinydashboard)
library (readxl)
 
```

```{r Global knitr Setup, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}

knitr::opts_chunk$set(
	#Chunk Code Options
	echo = FALSE, # Show the code in the output
	include = TRUE, # Show the results in the output.
	
	message = FALSE, # messages generated from code are not shown.
	warning = FALSE, # warnings generated from code are not shown.
	
	results = 'asis', # no additional formatting is applied to the results
	comment = NA, 
	
	# don't add the prompt character in R Code, i.e., 
	# don't stop to ask the user for execution
	prompt = FALSE, 	
	
	# # Plot Options - set the width and height (in inches) of all plots generated.
	# fig.height = 6,
	# fig.width = 8,
	# 
	# fig.retina only applies to html output 
	# controls the quality (dpi) of the image.
	fig.retina = 2,
	
	# default alignment of all figures.
	fig.align = 'center'
)

# round all decimals to 4 digits (after the decimal),
# unless specified otherwise in the local options. 
options(digits = 4)

```

```{r load data, include = FALSE, echo=FALSE}

unr <- read_xlsx("Aggregated_unr.xlsx")
educ_rate <- read_xls("education rate (Recession).xls")
deaths <-  read_xlsx("Deaths.xlsx")
ment_heal <- read_csv("mental_health.csv")
housing <- read_xlsx("Archived Zillow Data.xlsx")
per_capita <- read_xls("Personal_Income.xls")

``` 


# Unemployment


#### {}

<h2>Its impact on Select States.</h2>

## Column{.tabset}

### California
```{r}


g <- data.frame(
	year = c(2004, 2008, 2014),
	rate_unemployment = c(6.0, 5.0, 8.0),
	label=unr %>% select(`Period`) %>% unique()
	
)

unr %>% 
	filter(State =="California") %>% 
	ggplot(aes(x=year, y = `rate_unemployment`))+
	geom_col(aes(fill=`Period`) , alpha=0.6, position = "dodge")+
	geom_path(size=2)+
	geom_point(size=5, alpha=0.4)+
	geom_smooth(alpha=0.4)+
	geom_line(aes( group = `Period`), arrow=arrow(ends = "both", type = "closed"))+
	scale_fill_brewer(palette="Set1")+
  geom_label(
		data=g,
		size=2,
		aes(label = `Period`, fill=`Period`)
		)+
	labs(
      title = "California Unemployment Rate (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Unemployment Rate(%)")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18))

remove(g)
```

### Florida
```{r}
h <- data.frame(
	year = c(2004, 2008, 2012),
	rate_unemployment = c(3, 5, 7),
	label=unr %>% select(`Period`) %>% unique()
	
)

unr %>% 
	filter(State =="Florida") %>% 
	ggplot(aes(x=year, y = `rate_unemployment`))+
	geom_col(aes(fill=`Period`) , alpha=0.6, position = "dodge")+
	geom_path(size=2)+
	geom_point(size=5, alpha=0.4)+
	geom_smooth(alpha=0.4)+
	geom_line(aes( group = `Period`), arrow=arrow(ends = "both", type = "closed"))+
	scale_fill_brewer(palette="Pastel2")+
      	geom_label(
		data=h,
		size=2,
		aes(label = `Period`, fill=`Period`)
		)+
	labs(
      title = "Forida Unemployment Rate (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Unemployment Rate(%)")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18)) 
remove(h)
```

### Massechusetts
```{r}

i <- data.frame(
	year = c(2004, 2008, 2012),
	rate_unemployment = c(2, 2.5, 3),
	label=unr %>% select(`Period`) %>% unique()
	
)

unr %>% 
	filter(State =="Massechusetts") %>% 
	ggplot(aes(x=year, y = `rate_unemployment`))+
	geom_col(aes(fill=`Period`) , alpha=0.6, position = "dodge")+
	geom_path(size=2)+
	geom_point(size=5, alpha=0.4)+
	geom_smooth(alpha=0.4)+
	geom_line(aes( group = `Period`), arrow=arrow(ends = "both", type = "closed"))+
	scale_fill_brewer(palette="Set3")+
      	geom_label(
		data=i,
		size=2,
		aes(label = `Period`, fill=`Period`)
		)+
	labs(
      title = "Massechusetts Unemployment Rate (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Unemployment Rate(%)")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18))

remove(i)
```

### Michigan
```{r}

j <- data.frame(
	year = c(2004, 2008, 2012),
	rate_unemployment = c(3, 5, 7),
	label=unr %>% select(`Period`) %>% unique()
	
)

unr %>% 
	filter(State =="Michigan") %>% 
	ggplot(aes(x=year, y = `rate_unemployment`))+
	geom_col(aes(fill=`Period`) , alpha=0.6, position = "dodge")+
	geom_path(size=2)+
	geom_point(size=5, alpha=0.4)+
	geom_smooth(alpha=0.4)+
	geom_line(aes( group = `Period`), arrow=arrow(ends = "both", type = "closed"))+
	scale_fill_brewer(palette="Accent")+
    	geom_label(
		data=j,
		size=2,
		aes(label = `Period`, fill=`Period`)
		)+
	labs(
      title = "Michigan Unemploymnet Rate (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Unemployment Rate(%)")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18))
remove(j) 
```

### North Carolina
```{r}
k <- data.frame(
	year = c(2004, 2008, 2012),
	rate_unemployment = c(3, 5, 7),
	label=unr %>% select(`Period`) %>% unique()
	
)

unr %>% 
	filter(State =="North Carolina") %>% 
	ggplot(aes(x=year, y = `rate_unemployment`))+
	geom_col(aes(fill=`Period`) , alpha=0.6, position = "dodge")+
	geom_path(size=2)+
	geom_point(size=5, alpha=0.4)+
	geom_smooth(alpha=0.4)+
	geom_line(aes( group = `Period`), arrow=arrow(ends = "both", type = "closed"))+
	scale_fill_brewer(palette="Paired")+
      	geom_label(
		data=k,
		size=2,
		aes(label = `Period`, fill=`Period`)
		)+
	labs(
      title = "North Carolina Unemployment Rate (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Unemployment Rate(%)")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18))
remove(k)
```


### Comparing on the same graph 
```{r}

c <- unr %>% 
	group_by(State) %>% 
	ggplot(aes(x=year, y = `rate_unemployment`, (color='State')))+
	geom_point()+
	geom_smooth	(aes(color=State))+
	scale_fill_brewer(palette="Set1")+ # Use scale brewer to color the plot.
	labs( # Add Title, subtitle and labels to the axes.
	    title = "Education Rate by State",
	    subtitle = "Dynamic Axes - highlights difference",
	    fill="State",
	    x= "Year",
	    y = "Employment Rate")+
	theme(legend.position="")



d <- unr %>% 
	group_by(State) %>% 
	ggplot(aes(x=year, y = `rate_unemployment`, (color='State')))+
	geom_point()+
	geom_smooth	(aes(color=State))+
	scale_y_continuous(
		limits = c(0, 25), 
		labels = function(x) paste0(x, "%")
		) +
	scale_fill_brewer(palette="Set1")+ # Use scale brewer to color the plot.
	labs( # Add Title, subtitle and labels to the axes.
	    title = "Employment Rate by State",
	    subtitle = "Standard Axes - shows magnitude",
	    fill="State",
	    x= "Year",
	    y = "")+
	theme(legend.position="bottom")

grid.arrange(c, d, ncol=2)
remove(c, d)
```

# Unemployment Rate (Interactive)


####  {}
<h2>Its impact on California and Michigan.</h2>
Add your analysis / findings / Annotations here.

## Column{data-width=200}
### Select States and Period
```{r}

selectInput(
	inputId = "StateSelected",
	label = "Please select the state",
	choices = 
		unr %>% 
		select(State) %>% 
		unique()
)

```

```{r SelectPeriod}

selectInput(
	inputId = "PeriodSelected",
	label = "Please select the Period",
	choices = 
		unr %>% 
		select(Period) %>% 
		unique()
)

```

## Column{}
### Plotting Unemployment Rate for Selected State and Period
```{r}

renderPlot({

	unr %>%
	filter(State ==input$StateSelected) %>%
	filter(Period ==input$PeriodSelected) %>% 
	ggplot(aes(x=year, y = `rate_unemployment`))+
	geom_col(aes(fill=`Period`) , alpha=0.6, position = "dodge")+
	geom_path(size=2)+
	geom_point(size=5)+
	geom_smooth(alpha=0.4)+
	geom_line(
		aes( group = `Period`),
		arrow=arrow(ends = "both", type = "closed"))+
	scale_fill_brewer(palette="Paired")+
	labs(
      title = input$StateSelected,
      subtitle = "Unemployment Rate (%)",
      x= "Year",
      y = "Unemployment Rate(%)")+
	theme(
		plot.title =
			element_text(margin = margin(t = 10, b = 10), size=18),
		legend.position = 'top'
		)

})

```

### Table
```{r}

renderDT(
	unr %>% 
	select(Period, State, year,
		   'Unemployment(%)' = `rate_unemployment`) %>% 
	filter(State ==input$StateSelected) %>% 
	filter(Period ==input$PeriodSelected)
)

```

## Column{}
### All Years for Selected States
```{r}

renderPlot({

	unr %>%
	filter(State ==input$StateSelected) %>%
	ggplot(aes(x=year, y = `rate_unemployment`))+
	geom_col(aes(fill=`Period`) , alpha=0.6, position = "dodge")+
	geom_path(size=2)+
	geom_point(size=5)+
	geom_smooth(alpha=0.4)+
	geom_line(
		aes( group = `Period`),
		arrow=arrow(ends = "both", type = "closed"))+
	scale_fill_brewer(palette="Paired")+
	labs(
      title = input$StateSelected,
      subtitle = "Unemployment Rates (%)",
      x= "Year",
      y = "Unemployment(%)")+
	theme(
		plot.title =
			element_text(margin = margin(t = 10, b = 10), size=18),
		legend.position = 'top'
		)

})

```

# Per Capita Income
####  {}
<h2>Its impact on Select States.</h2>
Note that you can add regular HTML in your RMD and the browser will render the HTML. 
It is important to note that the presentation of the data and the axes on the charts makes a significant impact on the visual being presented and the interpretation that the user makes.

## Column{.tabset}

### California
```{r Per Capita Income}

ca_l <- data.frame(
	Year = c(2004, 2008, 2014),
	per_capita_income = c(37000, 43000, 51000),
	label=per_capita %>% select(`Period`) %>% unique()
)


per_capita %>% 
	filter(State == "California") %>% 
  ggplot(aes(x=Year, y = per_capita_income ))+
geom_point(aes(colour = `Period`), size = 5)+
  geom_smooth(alpha=0.3, se=TRUE, colour = "magenta 4")+
  scale_color_manual(values = c("Pre-recession" = "thistle 3", "Recession" = "deeppink", "Post-recession" = "Slateblue 1"))+
    geom_label(
		data=ca_l,
		size=2,
		aes(label = `Period`)
		)+
	labs(
      title = "Califronia Per Capita Income", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Per Capita Income")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18))
remove(ca_l)

```

### Florida
```{r}

fl_l <- data.frame(
	Year = c(2004, 2008, 2014),
	per_capita_income = c(36000, 40000, 44000),
	label=per_capita %>% select(`Period`) %>% unique()
)

per_capita %>% 
	filter(State == "Florida") %>% 
  ggplot(aes(x=Year, y = per_capita_income))+
geom_point(aes(colour = `Period`), size = 5)+
  geom_smooth(alpha=0.3, se=TRUE, colour = "Red 2")+
    scale_color_manual(values = c("Pre-recession" = "lightskyblue 2", "Recession" = "turquoise 2", "Post-recession" = "blue 2"))+
      geom_label(
		data=fl_l,
		size=2,
		aes(label = `Period`)
		)+
	labs(
      title = "Florida Per Capita Income (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Per Capita Income")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18))
remove(fl_l)

```

### Massachusetts
```{r}

m_l <- data.frame(
	Year = c(2004, 2008, 2014),
	per_capita_income = c(42000, 53000, 59000),
	label=per_capita %>% select(`Period`) %>% unique()
)

per_capita %>% 
	filter(State == "Massachusetts") %>% 
  ggplot(aes(x=Year, y = per_capita_income))+
geom_point(aes(colour = `Period`), size = 5)+
  geom_smooth(alpha=0.3, se=TRUE, colour = "Orange 2")+
    scale_color_manual(values = c("Pre-recession" = "darkolivegreen 2", "Recession" = "springgreen 2", "Post-recession" = "green 4"))+
  geom_label(
		data=m_l,
		size=2,
		aes(label = `Period`)
		)+
	labs(
      title = "Massachusetts Per Capita Income (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Per Capita Income")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18))
remove(m_l)

```

### North Carolina
```{r}

n_l <- data.frame(
	Year = c(2004, 2008, 2014),
	per_capita_income = c(30000, 37000, 41000),
	label=per_capita %>% select(`Period`) %>% unique()
)

per_capita %>% 
	filter(State == "North Carolina") %>% 
  ggplot(aes(x=Year, y = per_capita_income))+
geom_point(aes(colour = `Period`), size = 5)+
  geom_smooth(alpha=0.3, se=TRUE, colour = "Yellow 2")+
    scale_color_manual(values = c("Pre-recession" = "blueviolet", "Recession" = "violetred 2", "Post-recession" = "plum 2"))+
    geom_label(
		data=n_l,
		size=2,
		aes(label = `Period`)
		)+
	labs(
      title = "North Carolina Per Capita Income (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Per Capita Income")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18))
remove(n_l)

```

### Compare by states
```{r}

per_capita %>% 
  ggplot(aes(x=Year, y = per_capita_income))+
geom_point(aes(colour = `Period`), size = 5)+
  geom_smooth(alpha=0.3, se=TRUE, colour = "Orange 3")+
      scale_color_manual(values = c("Pre-recession" = "red", "Recession" = "blue", "Post-recession" = "green 4"))+
	labs(
      title = " Per Capita Income By States (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Per Capita Income")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18))+
  facet_wrap(~ State)

```

# Education

####  {}
<h2>Its impact on Select States.</h2>
Note that you can add regular HTML in your RMD and the browser will render the HTML. 
It is important to note that the presentation of the data and the axes on the charts makes a significant impact on the visual being presented and the interpretation that the user makes.

## Column{.tabset}

### California
```{r Education}

cali <- data.frame(
	year = c(2006, 2008, 2014),
	education_rate = c(29, 29, 31),
	label=educ_rate %>% select(`Period`) %>% unique()
	
)


educ_rate %>% 
	filter(State =="California") %>% 
	ggplot(aes(x=year, y = `education_rate`))+
	geom_col(aes(fill=`Period`) , alpha=0.6, position = "dodge")+
	geom_path(size=2)+
	geom_point(size=5, alpha=0.4)+
	geom_smooth(alpha=0.4)+
	geom_line(aes( group = `Period`), arrow=arrow(ends = "both", type = "closed"))+
	scale_fill_brewer(palette="Set3")+
      	geom_label(
		data=cali,
		size=2,
		aes(label = `Period`, fill=`Period`)
		)+
	labs(
      title = "California Education Rates (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Education Rate(%)")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18)) #top and bottom margins and font size
remove(cali)

```

### Florida 
```{r}
flori <- data.frame(
	year = c(2006, 2008, 2014),
	education_rate = c(25, 25, 27),
	label=educ_rate %>% select(`Period`) %>% unique()
	
)

educ_rate %>% 
	filter(State =="Florida") %>% 
	ggplot(aes(x=year, y = `education_rate`))+
	geom_col(aes(fill=`Period`) , alpha=0.6, position = "dodge")+
	geom_path(size=2)+
	geom_point(size=5, alpha=0.4)+
	geom_smooth(alpha=0.4)+
	geom_line(
		aes( group = `Period`), 
		arrow=arrow(ends = "both", type = "closed"))+
	scale_fill_brewer(palette="PRGn")+
        	geom_label(
		data=flori,
		size=2,
		aes(label = `Period`, fill=`Period`)
		)+
	labs(
      title = "Florida Education Rate (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Education Rate (%)")+
	theme(
		plot.title = 
			element_text(margin = margin(t = 5, b = 5), size=14),
		legend.position="bottom"
		)
remove(flori)

```


### North Carolina
```{r}
nocar <- data.frame(
	year = c(2006, 2008, 2014),
	education_rate = c(25, 25, 25),
	label=educ_rate %>% select(`Period`) %>% unique()
	
)

educ_rate %>% 
	filter(State =="North Carolina") %>% 
	ggplot(aes(x=year, y = `education_rate`))+
	geom_col(aes(fill=`Period`) , alpha=0.6, position = "dodge")+
	geom_path(size=2)+
	geom_point(size=5, alpha=0.4)+
	geom_smooth(alpha=0.4)+
	geom_line(aes( group = `Period`), arrow=arrow(ends = "both", type = "closed"))+
	scale_fill_brewer(palette="Paired")+
        	geom_label(
		data=nocar,
		size=2,
		aes(label = `Period`, fill=`Period`)
		)+
	labs(
      title = "North carolina Education Rates (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Education Rate(%)")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18)) #top and bottom margins and font size
remove(nocar)

```

### Massachusetts
```{r}
ma <- data.frame(
	year = c(2006, 2008, 2014),
	education_rate = c(25, 25, 25),
	label=educ_rate %>% select(`Period`) %>% unique()
	
)

educ_rate %>% 
	filter(State =="Massachusetts") %>% 
	ggplot(aes(x=year, y = `education_rate`))+
	geom_col(aes(fill=`Period`) , alpha=0.6, position = "dodge")+
	geom_path(size=2)+
	geom_point(size=5, alpha=0.4)+
	geom_smooth(alpha=0.4)+
	geom_line(aes( group = `Period`), arrow=arrow(ends = "both", type = "closed"))+
	scale_fill_brewer(palette="Set1")+
        	geom_label(
		data=ma,
		size=2,
		aes(label = `Period`, fill=`Period`)
		)+
	labs(
      title = "Massachusetts Education Rates (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Education Rate(%)")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18)) #top and bottom margins and font size
remove(ma)

```


### Comparing
```{r}

a<-educ_rate %>% 
	group_by(State) %>% 
	ggplot(aes(x=year, y = `education_rate`, (color='State')))+
	geom_point()+
	geom_smooth	(aes(color=State))+
	scale_fill_brewer(palette="PRGn")+ # Use scale brewer to color the plot.
	labs( # Add Title, subtitle and labels to the axes.
	    title = "Education Rate by State",
	    subtitle = "Dynamic Axes - highlights difference",
	    fill="State",
	    x= "Year",
	    y = "Education Rate")+
	theme(legend.position="")



b<-educ_rate %>% 
	group_by(State) %>% 
	ggplot(aes(x=year, y = `education_rate`, (color='State')))+
	geom_point()+
	geom_smooth	(aes(color=State))+
	scale_y_continuous(
		limits = c(0, 100), 
		labels = function(x) paste0(x, "%")
		) +
	scale_fill_brewer(palette="PRGn")+ # Use scale brewer to color the plot.
	labs( # Add Title, subtitle and labels to the axes.
	    title = "Education Rate by State",
	    subtitle = "Standard Axes - shows magnitude",
	    fill="State",
	    x= "Year",
	    y = "")+
	theme(legend.position="bottom")

grid.arrange(a, b, ncol=2)
remove(a, b)
```



# Select Vars

#### 
<h2>Its impact on States</h2>
Add your analysis / findings / Annotations here.

## Column{data-width=150}
### Select States and Period
```{r TheSelectPeriod}

selectInput(
	inputId = "ThePeriodSelected",
	label = "Please select the Period",
	choices = 
		educ_rate %>% 
		select(Period) %>% 
		unique()
)

```

```{r State}

selectInput(
	inputId = "TheStateSelected",
	label = "Please select the state",
	choices = 
		educ_rate %>% 
		select(State) %>% 
		unique()
)

```

```{r SelectVars}


selectInput(
	inputId = "VariablesSelected",
	label = "Please select the Variable",
	choices = c(educ_rate %>% select('year', 'education_rate') %>% names())
)


```


## Column{}
### Plotting Home Ownersip for Selected State and Period
```{r}

renderPlot({

myData <- educ_rate %>%
	filter(State ==input$TheStateSelected) %>%
	select(Period, State, year,input$VariablesSelected)

ggplot(myData,
		aes_string( 
			x= "year",
			y = input$VariablesSelected))+
	geom_col(aes(fill=`Period`) , alpha=0.6, position = "dodge")+
	geom_point(size=5)+
	geom_smooth(alpha=0.4, se=TRUE)+
	scale_fill_brewer(palette="Paired")+
	labs(
      title = input$TheStateSelected,
      subtitle = input$VariablesSelected,
      x= "Year",
      y = input$VariablesSelected)+
	theme(
		plot.title = element_text(
			margin = margin(t = 10, b = 10), size=18),
		legend.position = 'right'
		)

})

```



### Table{}
```{r}

renderDT(
	educ_rate %>% 
	filter(State ==input$TheStateSelected) %>% 
	filter(Period ==input$ThePeriodSelected) %>% 
	select(Period, State, year,input$VariablesSelected)
)

```


# Compare States

#### 
<h2>Impact on States</h2>
Add your analysis / findings / Annotations here.

## Column{data-width=150}
### Select Period and Variables
```{r Clean Var Names}


```


```{r SelectVars2}



selectInput(
	inputId = "VariablesSelected1",
	label = "Please select the Variable",
	choices = c(educ_rate %>% select('year', 'education_rate') %>% names())
)


```

```{r State1}

selectInput(
	inputId = "TheStateSelected1",
	label = "Please select the first state",
	choices = 
	educ_rate %>% 
		select(State) %>% 
		unique()
)

```

```{r State2}

selectInput(
	inputId = "TheStateSelected2",
	label = "Please select the second state",
	choices = 
		educ_rate %>% 
		arrange(desc(State)) %>% 
		select(State) %>% 
		unique()
)

```

## Column{}

### Compare State 1

```{r}

renderPlot({

educ_rate %>%
	filter(State ==input$TheStateSelected1) %>%
	select(Period, State, year,input$VariablesSelected1) %>% 
	ggplot(aes_string( 
			x= "year",
			y = input$VariablesSelected1))+
	geom_col(aes(fill=`Period`) , alpha=0.6, position = "dodge")+
	geom_point(size=5)+
	geom_smooth(alpha=0.4, se=TRUE)+
	scale_fill_brewer(palette="Paired")+
	labs(
      title = input$TheStateSelected1,
      subtitle = input$VariablesSelected1,
      x= "Year",
      y = input$VariablesSelected1)+
	theme(
		plot.title = element_text(
			margin = margin(t = 10, b = 10), size=18),
		legend.position = 'top'
		)

})

```


### Compare State 2

```{r}

renderPlot({

educ_rate %>%
	filter(State ==input$TheStateSelected2) %>%
	select(Period, State, year,input$VariablesSelected1) %>% 
	ggplot(aes_string( 
			x= "year",
			y = input$VariablesSelected1))+
	geom_col(aes(fill=`Period`) , alpha=0.6, position = "dodge")+
	geom_point(size=5)+
	geom_smooth(alpha=0.4, se=TRUE)+
	scale_fill_brewer(palette="Paired")+
	labs(
      title = input$TheStateSelected2,
      subtitle = input$VariablesSelected1,
      x= "Year",
      y = input$VariablesSelected1)+
	theme(
		plot.title = element_text(
			margin = margin(t = 10, b = 10), size=18),
		legend.position = 'top'
		)

})

```


# Housing Maps {data-orientation="rows"}

## Row {data-height="100"}

Homeownership Rates - Extend to other indicators - Add Text to compare rates.

## Row {data-height="600"}

### Pre-Recession
```{r maps Pre-recession}
us_map <- map_data('state')

HousingData <- housing %>%
	filter(Period == "Pre-recession") %>% 
	group_by(State) %>% 
	summarise(
		MeanIndex = mean(`index`)
	)

#Convert to lower case.
HousingData$region <- gsub('_',' ',tolower(HousingData$State))

merged_data <- left_join(us_map, HousingData[, c('region', 'MeanIndex')], by='region')

 
merged_data %>% 
	ggplot()+
	geom_polygon(aes(x=long, y=lat, group=group, fill=MeanIndex))+
	theme_void()+
   labs(title = 'Mean Housing Index', 
        subtitle = 'Pre-Recession')
```

### Recession
```{r maps Recession}

# Recession Time Plot
#---------------------------------
HousingData <- housing %>%
	filter(Period == "Recession") %>% 
	group_by(State) %>% 
	summarise(
		MeanIndex = mean(`index`)
	)

#Convert to lower case.
HousingData$region <- gsub('_',' ',tolower(HousingData$State))

merged_data <- left_join(us_map, HousingData[, c('region', 'MeanIndex')], by='region')

 
merged_data %>% 
	ggplot()+
	geom_polygon(aes(x=long, y=lat, group=group, fill=MeanIndex))+
	theme_void()+
   labs(title = 'Mean Housing Index', 
        subtitle = 'Recession')


```

### Post-Recession

```{r maps Post-Recession}

# Post-Recession Time Plot
#---------------------------------
HousingData <- housing %>%
	filter(Period == "Post-recession") %>% 
	group_by(State) %>% 
	summarise(
		MeanIndex = mean(`index`)
	)

#Convert to lower case.
HousingData$region <- gsub('_',' ',tolower(HousingData$State))

merged_data <- left_join(us_map, HousingData[, c('region', 'MeanIndex')], by='region')

 
merged_data %>% 
	ggplot()+
	geom_polygon(aes(x=long, y=lat, group=group, fill=MeanIndex))+
	theme_void()+
   labs(title = 'Mean Housing Index', 
        subtitle = 'Post-recession')

remove(HousingData, merged_data)
```

## Row
### Table
```{r}

renderDT(
	housing %>% 
	select(Period, State, Date,
		   'Index' = `index`)

)

```

# Death Rates

####  {}
<h2>Its impact on Select States.</h2>
Note that you can add regular HTML in your RMD and the browser will render the HTML. 
It is important to note that the presentation of the data and the axes on the charts makes a significant impact on the visual being presented and the interpretation that the user makes.

## Column{.tabset}

### California 
```{r}
deaths %>% filter(State=="California") %>% 
ggplot( aes(x = Year, y = `death_rate`)) +
  geom_bar(stat='identity', aes(fill = `Period`))+
  ggtitle("California Death Rate Due To Health Impact")+
    	scale_fill_brewer(palette="Accent")+
	labs(
      title = "California Death Rates By Cause (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Death Rate(%)")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18))+
  facet_wrap(~ Cause)
```

### Florida
```{r}
deaths %>% filter(State=="Florida") %>% 
ggplot(aes(x = Year, y = `death_rate`)) +
  geom_bar(stat='identity', aes(fill = `Period`))+
  ggtitle("Florida Death Rate Due To Health Impact")+
    	scale_fill_brewer(palette="Dark2")+
	labs(
      title = "Florida Death Rates By Cause (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Death Rate(%)")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18))+
  facet_wrap(~ Cause)
```

### Massachusetts
```{r}
deaths %>% filter(State=="Massachusetts") %>% 
ggplot(aes(x = Year, y = `death_rate`)) +
  geom_bar(stat='identity', aes(fill = `Period`))+
  ggtitle("Massachusetts Death Rate Due To Health Impact")+
    	scale_fill_brewer(palette="Set3")+
	labs(
      title = "Massachusetts Death Rates By Cause (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Death Rate(%)")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18))+
  facet_wrap(~ Cause)

```

### Michigan
```{r}
deaths %>% filter(State=="Michigan") %>% 
ggplot(aes(x = Year, y = `death_rate`)) +
  geom_bar(stat='identity', aes(fill = `Period`))+
  ggtitle("Michigan Death Rate Due To Health Impact")+
    	scale_fill_brewer(palette="Set1")+
	labs(
      title = "Michigan Death Rates By Cause (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Death Rate(%)")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18))+
  facet_wrap(~ Cause)
```

### North Carolina
```{r}
deaths %>% filter(State=="North Carolina") %>% 
ggplot(aes(x = Year, y = `death_rate`)) +
  geom_bar(stat='identity', aes(fill = `Period`))+
  ggtitle("North Carolina Death Rate Due To Health Impact")+
  	scale_fill_brewer(palette="Paired")+
	labs(
      title = "North Carolina Death Rates By Cause (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Death Rate(%)")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18))+
  facet_wrap(~ Cause)
```

### Cause Of Deaths
```{r}

ggplot(deaths, aes(x = Year, y = `death_rate`)) +
  geom_bar(stat='identity', aes(fill = `Period`))+
  ggtitle("Death Rates By Cause")+
  facet_wrap(~ Cause)
```


# Mental Health Rate

####  {}
<h2>Its impact on Select States.</h2>
Note that you can add regular HTML in your RMD and the browser will render the HTML. 
It is important to note that the presentation of the data and the axes on the charts makes a significant impact on the visual being presented and the interpretation that the user makes.

## Column{.tabset}

### California
```{r MentalHealthRate}

Cal_lab <- data.frame(
	Year = c(2004, 2008, 2014),
	depressive_disorder = c(4.000, 3.500, 3.500),
	label=ment_heal %>% select(`Period`) %>% unique()
)

ment_heal %>% 
	filter(State =="California") %>% 
	ggplot(aes(x=Year, y = `depressive_disorder`))+
	geom_line( alpha=0.9, position = "identity",colour = "dark blue", size=1)+
	geom_point(color="Orange", size=3, alpha=0.9)+
    geom_label(
		data=Cal_lab,
		size=2,
		aes(label = `Period`)
		)+
	labs(
      title = "California Depressive Disorder Rate (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Depressive Disorder (%)")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18))
remove(Cal_lab)

```

### Florida
```{r}

flo_lab <- data.frame(
	Year = c(2004, 2008, 2014),
	depressive_disorder = c(3.350, 3.370, 3.500),
	label=ment_heal %>% select(`Period`) %>% unique()
)

ment_heal %>% 
	filter(State =="Florida") %>% 
	ggplot(aes(x=Year, y = `depressive_disorder`))+
	geom_line( alpha=0.9, position = "identity", colour = "maroon", size=1)+
	geom_point(color="Yellow", size=3, alpha=0.9)+
  geom_label(
		data=flo_lab,
		size=2,
		aes(label = `Period`)
		)+
	labs(
      title = "Florida Depressive Disorder Rate (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Depressive Disorder (%)")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18))
remove(flo_lab)
```

### Massachusetts
```{r}

ma_lab <- data.frame(
	Year = c(2004, 2008, 2014),
	depressive_disorder = c(5.130, 5.140, 5.000),
	label=ment_heal %>% select(`Period`) %>% unique()
)

ment_heal %>% 
	filter(State =="Massachusetts") %>% 
	ggplot(aes(x=Year, y = `depressive_disorder`))+
	geom_line( alpha=0.9, position = "identity", colour = "dark green", size=1)+
	geom_point(color="Red", size=3, alpha=0.9)+
    geom_label(
		data=ma_lab,
		size=2,
		aes(label = `Period`)
		)+
	labs(
      title = "Massachusetts Depressive Disorder Rate (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Depressive Disorder (%)")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18))
remove(ma_lab)
```

### North Carolina
```{r}

nc_lab <- data.frame(
	Year = c(2004, 2008, 2014),
	depressive_disorder = c(4.200, 4.430, 4.400),
	label=ment_heal %>% select(`Period`) %>% unique()
)


ment_heal %>% 
	filter(State =="North Carolina") %>% 
	ggplot(aes(x=Year, y = `depressive_disorder`))+
	geom_line( alpha=0.9, position = "identity", colour = "violet", size=1)+
	geom_point(color="Blue", size=3, alpha=0.9)+
      geom_label(
		data=nc_lab,
		size=2,
		aes(label = `Period`)
		)+
	labs(
      title = "North Carolina Depressive Disorder Rate (%)", 
      subtitle = "Note the Y-Axes",
      x= "Year",
      y = "Depressive Disorder (%)")+
	theme(plot.title = element_text(margin = margin(t = 10, b = 10), size=18))
remove(nc_lab)
```


### Comparing on the same graph 
```{r}


e<-ment_heal %>% 
	group_by(State) %>% 
	ggplot(aes(x=Year, y = `depressive_disorder`, (color='State')))+
	geom_point()+
	geom_smooth	(aes(color=State))+
	scale_fill_brewer(palette="Set1")+ # Use scale brewer to color the plot.
	labs( 
	    title = "Depressive Disorder by State",
	    subtitle = "Dynamic Axes - highlights difference",
	    fill="State",
	    x= "Year",
	    y = "Depressive Disorder Rate")+
	theme(legend.position="")



f<- ment_heal %>% 
	group_by(State) %>% 
	ggplot(aes(x=Year, y = `depressive_disorder`, (color='State')))+
	geom_point()+
	geom_smooth	(aes(color=State))+
	scale_y_continuous(
		limits = c(0, 100), 
		labels = function(x) paste0(x, "%")
		) +
	scale_fill_brewer(palette="Set1")+ # Use scale brewer to color the plot.
	labs( # Add Title, subtitle and labels to the axes.
	    title = "Depressive Disorder by State",
	    subtitle = "Standard Axes - shows magnitude",
	    fill="State",
	    x= "Year",
	    y = "")+
	theme(legend.position="bottom")

grid.arrange(e, f, ncol=2)
remove(e, f)
```

